---
title: "Prepare data"
author: "Stefano Coretta"
format: html
editor: visual
---

## Attach packages

```{r pkgs}
library(here)
library(tidyverse)
library(faux)
```

## Glottolog data

The Glottolog data is available in the \[glottologR\](https://github.com/stefanocoretta/glottologR) package.

## Politeness

```{r}
polite <- read_csv("https://raw.githubusercontent.com/learnB4SS/learnB4SS/main/data-raw/polite.csv") %>%
  arrange(index) %>%
  select(-index, -years_in_ger)

usethis::use_data(polite, overwrite = TRUE)
```

## Vowel formants

```{r}
f_ita_pol <- read_csv("https://raw.githubusercontent.com/stefanocoretta/coretta2018itapol/master/data-raw/aggregated/formants-series.csv") %>%
  select(everything(), gender = sex)

usethis::use_data(f_ita_pol, overwrite = TRUE)
```

```{r}
f_mid <- f_ita_pol %>%
  filter(time == 5) %>%
  select(-time)

dur_ita_pol <- read_csv("https://raw.githubusercontent.com/stefanocoretta/coretta2018itapol/master/data-raw/aggregated/token-measures.csv") %>%
  select(everything(), gender = sex) %>%
  left_join(y = f_mid)

usethis::use_data(dur_ita_pol, overwrite = TRUE)
```

## Emotional Valence

```{r}
set.seed(74201)
emotion <- read_csv("https://osf.io/u47fk/download") %>%
  mutate(
    valence = Val - 5,
    valence_z = as.numeric(scale(Val))
  ) %>%
  sample_n(1000, weight = pnorm(valence_z)) %>%
  select(word = Word, valence)

usethis::use_data(emotion, overwrite = TRUE)
```

## Morphology and dialectology in the Linguistic Survey of Scotland

The [Morphology and dialectology in the Linguistic Survey of Scotland](https://osf.io/4bp2y) contains a portion of the morphological questionnaire returns collected for the Linguistic Survey of Scotland (Gaelic). The data have been transcribed and coded from questionnaire slips by Will Lamb.

Iosad, Pavel & William Lamb. 2020. *Morphology and dialectology in the Linguistic Survey of Scotland*. OSF. [osf.io/4bp2y](https://osf.io/4bp2y).

```{r}
gael_morph <- read_csv("https://osf.io/download/h2vub/")

usethis::use_data(gael_morph, overwrite = TRUE)
```

## Massive Auditory Lexical Decision 1.1

```{r}
download.file(
  "https://era.library.ualberta.ca/items/3344343b-2b4a-4b8c-af8e-8bd829c76472/view/eada02c5-6259-4225-8204-13ce86e1cd2d/MALD1_1_AllData.Rda",
  here("./data-raw/MALD1_1_AllData.Rda")
)

load(here("./data-raw/MALD1_1_AllData.Rda"))
set.seed(943874)

subj <- sample(unique(dat$Subject), 50)

mald_1_1 <- dat %>%
  filter(Subject %in% subj, RT > 0) %>%
  select(Subject, Item, IsWord, PhonLev, RT, ACC) %>%
  mutate(
    IsWord = factor(IsWord, levels = c("TRUE", "FALSE")),
    ACC = ifelse(ACC, "correct", "incorrect"),
    ACC = factor(ACC, levels = c("incorrect", "correct")),
    RT_log = log(RT)
  ) %>%
  group_by(Subject) %>%
  slice_sample(n = 100) %>%
  ungroup()

usethis::use_data(mald_1_1, overwrite = TRUE)
```

## 'Everywhere here can say this': The English locative impersonal

```{r}
everywhere_loc <- read_csv("https://osf.io/download/vaujp/?view_only=edba6a2255044ce6af6a519b85c9e65f") %>%
  select(-(Simple:Pass))

usethis::use_data(everywhere_loc, overwrite = TRUE)
```

## A Cross-Cultural Analysis of Early Prelinguistic Gesture Development and Its Relationship to Language Development

```{r}
gestures <- read_csv("https://raw.githubusercontent.com/stefanocoretta/infant-gestures/master/data/gestures.csv") %>%
  mutate(
    id = as.numeric(as.factor(paste(dyad, task, gesture)))
  )

usethis::use_data(gestures, overwrite = TRUE)
```

## Shallow morphological parsing

```{r}
shallow <- read_csv("https://raw.githubusercontent.com/stefanocoretta/quick-linear-models/main/04-fit-models/shallow-morph.csv") %>%
  filter(Critical_Filler == "Critical") %>%
  mutate(
    Relation_type = factor(Relation_type, levels = c("Unrelated", "Constituent", "NonConstituent")),
    accuracy = ifelse(ACC == 0, "incorrect", "correct")
  ) %>%
  select(-ACC, RT_log = logRT)

usethis::use_data(shallow, overwrite = TRUE)
write_csv(shallow, here("./data-raw/tutorials/week-5-tutorial-files/shallow.csv"))
```

## Italian EGG

```{r}
stimuli <- read_csv("https://raw.githubusercontent.com/stefanocoretta/coretta2018itaegg/master/data-raw/datasets/stimuli.csv")

ita_egg <- read_csv("https://raw.githubusercontent.com/stefanocoretta/coretta2018itaegg/master/data-raw/datasets/measurements.csv", na = "--undefined--") %>%
  full_join(y = stimuli) %>%
  mutate(
    v1_duration = (c2_ons - v1_ons) * 1000,
    c2_clos_duration = (c2_rel - c2_ons) * 1000,
    rel_voff = (c2_ons - c1_rel) * 1000,
    sent_duration = sentence_off - sentence_ons,
    speech_rate = 8 / sent_duration,
    speech_rate_c = speech_rate - mean(speech_rate, na.rm = TRUE),
    voice_duration = (voice_off - voice_ons) * 1000,
    vot = (voice_ons - c1_rel) * 1000,
    voi_clo = (voice_off - c2_ons) * 1000,
    rel_rel = (c2_rel - c1_rel) * 1000,
    vowel = factor(vowel, levels = c("i", "e", "a", "o", "u")),
    height = factor(height, levels = c("low", "mid-low", "mid-high", "high")),
    c1_place = factor(c1_place, levels = c("labial", "coronal", "velar")),
    c2_place = factor(c2_place, levels = c("labial", "coronal", "velar"))
  )

usethis::use_data(ita_egg, overwrite = TRUE)
```

## Formative assessments

### FT1: Data viz

```{r}
ft1_data <- read_csv("https://osf.io/download/vaujp/?view_only=edba6a2255044ce6af6a519b85c9e65f") %>%
  select(Subject, Condition, Restrictor, Value, Age, Education, Gender) %>%
  mutate(
    Restrictor = ifelse(Restrictor == 0, "non-restricted", "restricted"),
    Value = as.factor(Value)
  ) %>%
  filter(
    Condition %in% c("gold-good", "gold-bad", "simple", "adverb", "modal")
  )

usethis::use_data(ft1_data, overwrite = TRUE)
```

### FT2: Linear models

```{r}
download.file(
  "https://github.com/cran/languageR/raw/master/data/english.rda",
  here("data-raw/assessments/ft2/english.rda")
)

load(here("data-raw/assessments/ft2/english.rda"))

write_csv(english, here("data-raw/assessments/ft2/english.csv"))
```

### FT3: Data analysis

```{r}
download.file(
  "https://github.com/cran/languageR/raw/master/data/dative.rda",
  here("data-raw/assessments/ft3/dative.rda")
)

load(here("data-raw/assessments/ft3/dative.rda"))

dative <- dative %>%
  select(RealizationRec = RealizationOfRecipient, Verb, AnimacyOfRec, AnimacyOfTheme, LengthOfTheme)

write_csv(dative, here("data-raw/assessments/ft3/dative.csv"))
```

## Summative assessment

```{r}
load(here("data/dur_ita_pol.rda"))

data_e1_1 <- dur_ita_pol %>%
  select(
    speaker, f0_midpoint = f0, speech_rate, vowel, condition = c2_phonation, group = c2_place
  ) %>%
  drop_na()

write_csv(data_e1_1, here("data-raw/assessments/s1/data_e1_1.csv"))

n <- 400
language <- rep(c("Tagalog", "Maori"), each = n/2)
environment <- rep(c("quiet", "noisy"), length.out = n)
age <- rep(c("old", "old", "young", "young"), length.out = n)

RT_ms <- exp(rnorm(n, 5, 1) +
  (as.numeric(as.factor(language)) - 1) * rnorm(n, 0.5, 2) +
  (as.numeric(as.factor(environment)) - 1) * rnorm(n, -0.3, 1) +
  (as.numeric(as.factor(age)) - 1) * rnorm(n, 2, 1))

data_e1_2 <- tibble(
  RT_ms, language, environment, age
) %>%
  filter(RT_ms < 2000)

write_csv(data_e1_2, here("data-raw/assessments/s1/data_e1_2.csv"))

n <- 400
trial <- rep(1:12, length.out = n)
subject <- rep(1:9, each = 48)[1:400]
condition <- rep(rep(c("easy", "difficult"), each = 12), length.out = n)
priming <- rep(rep(c("high_freq", "low_freq"), each = 24), length.out = n)

accuracy <- sample(
  c("correct", "incorrect"), 400, replace = TRUE,
  prob = for (i in 1:400) {
      if (condition[i] == "easy") {
      c(0.8, 0.2)
    } else {
      c(0.1, 0.9)
    }
  }
)

data_e1_3 <- tibble(
  subject, trial, condition, priming, accuracy
)

write_csv(data_e1_3, here("data-raw/assessments/s1/data_e1_3.csv"))
```

```{r}

```
