---
title: "Prepare data"
author: "Stefano Coretta"
format: html
editor: visual
---

## Attach packages

```{r pkgs}
library(here)
library(tidyverse)
```

## Glottolog data

The Glottolog data is available in the \[glottologR\](https://github.com/stefanocoretta/glottologR) package.

## Politeness

```{r}
polite <- read_csv("https://raw.githubusercontent.com/learnB4SS/learnB4SS/main/data-raw/polite.csv") %>%
  arrange(index) %>%
  select(-index, -years_in_ger)

usethis::use_data(polite, overwrite = TRUE)
```

## Vowel formants

```{r}
f_ita_pol <- read_csv("https://raw.githubusercontent.com/stefanocoretta/coretta2018itapol/master/data-raw/aggregated/formants-series.csv") %>%
  select(everything(), gender = sex)

usethis::use_data(f_ita_pol, overwrite = TRUE)
```

## Emotional Valence

```{r}
set.seed(74201)
emotion <- read_csv("https://osf.io/u47fk/download") %>%
  mutate(
    valence = Val - 5,
    valence_z = as.numeric(scale(Val))
  ) %>%
  sample_n(1000, weight = pnorm(valence_z)) %>%
  select(word = Word, valence)

usethis::use_data(emotion, overwrite = TRUE)
```

## Morphology and dialectology in the Linguistic Survey of Scotland

The [Morphology and dialectology in the Linguistic Survey of Scotland](https://osf.io/4bp2y) contains a portion of the morphological questionnaire returns collected for the Linguistic Survey of Scotland (Gaelic). The data have been transcribed and coded from questionnaire slips by Will Lamb.

Iosad, Pavel & William Lamb. 2020. *Morphology and dialectology in the Linguistic Survey of Scotland*. OSF. [osf.io/4bp2y](https://osf.io/4bp2y).

```{r}
gael_morph <- read_csv("https://osf.io/download/h2vub/")

usethis::use_data(gael_morph, overwrite = TRUE)
```

