---
title: "Prepare data"
author: "Stefano Coretta"
format: html
editor: visual
---

## Attach packages

```{r pkgs}
library(here)
library(tidyverse)
```

## Glottolog data

The Glottolog data is available in the \[glottologR\](https://github.com/stefanocoretta/glottologR) package.

## Politeness

```{r}
polite <- read_csv("https://raw.githubusercontent.com/learnB4SS/learnB4SS/main/data-raw/polite.csv") %>%
  arrange(index) %>%
  select(-index, -years_in_ger)

usethis::use_data(polite, overwrite = TRUE)
```

## Vowel formants

```{r}
f_ita_pol <- read_csv("https://raw.githubusercontent.com/stefanocoretta/coretta2018itapol/master/data-raw/aggregated/formants-series.csv") %>%
  select(everything(), gender = sex)

usethis::use_data(f_ita_pol, overwrite = TRUE)
```

```{r}
f_mid <- f_ita_pol %>%
  filter(time == 5) %>%
  select(-time)

dur_ita_pol <- read_csv("https://raw.githubusercontent.com/stefanocoretta/coretta2018itapol/master/data-raw/aggregated/token-measures.csv") %>%
  select(everything(), gender = sex) %>%
  left_join(y = f_mid)

usethis::use_data(dur_ita_pol, overwrite = TRUE)
```

## Emotional Valence

```{r}
set.seed(74201)
emotion <- read_csv("https://osf.io/u47fk/download") %>%
  mutate(
    valence = Val - 5,
    valence_z = as.numeric(scale(Val))
  ) %>%
  sample_n(1000, weight = pnorm(valence_z)) %>%
  select(word = Word, valence)

usethis::use_data(emotion, overwrite = TRUE)
```

## Morphology and dialectology in the Linguistic Survey of Scotland

The [Morphology and dialectology in the Linguistic Survey of Scotland](https://osf.io/4bp2y) contains a portion of the morphological questionnaire returns collected for the Linguistic Survey of Scotland (Gaelic). The data have been transcribed and coded from questionnaire slips by Will Lamb.

Iosad, Pavel & William Lamb. 2020. *Morphology and dialectology in the Linguistic Survey of Scotland*. OSF. [osf.io/4bp2y](https://osf.io/4bp2y).

```{r}
gael_morph <- read_csv("https://osf.io/download/h2vub/")

usethis::use_data(gael_morph, overwrite = TRUE)
```

## Massive Auditory Lexical Decision 1.1

```{r}
download.file(
  "https://era.library.ualberta.ca/items/3344343b-2b4a-4b8c-af8e-8bd829c76472/view/eada02c5-6259-4225-8204-13ce86e1cd2d/MALD1_1_AllData.Rda",
  here("./data-raw/MALD1_1_AllData.Rda")
)

load(here("./data-raw/MALD1_1_AllData.Rda"))
set.seed(943874)

subj <- sample(unique(dat$Subject), 50)

mald_1_1 <- dat %>%
  filter(Subject %in% subj, RT > 0) %>%
  select(Subject, Item, IsWord, PhonLev, RT, ACC) %>%
  mutate(
    IsWord = factor(IsWord, levels = c("TRUE", "FALSE")),
    ACC = ifelse(ACC, "correct", "incorrect"),
    ACC = factor(ACC, levels = c("incorrect", "correct")),
    RT_log = log(RT)
  ) %>%
  group_by(Subject) %>%
  slice_sample(n = 100) %>%
  ungroup()

usethis::use_data(mald_1_1, overwrite = TRUE)
```

## 'Everywhere here can say this': The English locative impersonal

```{r}
everywhere_loc <- read_csv("https://osf.io/download/vaujp/?view_only=edba6a2255044ce6af6a519b85c9e65f") %>%
  select(-(Simple:Pass))

usethis::use_data(everywhere_loc, overwrite = TRUE)
```

## A Cross-Cultural Analysis of Early Prelinguistic Gesture Development and Its Relationship to Language Development

```{r}
gestures <- read_csv("https://raw.githubusercontent.com/stefanocoretta/infant-gestures/master/data/gestures.csv") %>%
  mutate(
    id = as.numeric(as.factor(paste(dyad, task, gesture)))
  )

usethis::use_data(gestures, overwrite = TRUE)
```

## Shallow morphological parsing

```{r}
shallow <- read_csv("https://raw.githubusercontent.com/stefanocoretta/quick-linear-models/main/04-fit-models/shallow-morph.csv") %>%
  filter(Critical_Filler == "Critical") %>%
  mutate(
    Relation_type = factor(Relation_type, levels = c("Unrelated", "Constituent", "NonConstituent")),
    accuracy = ifelse(ACC == 0, "incorrect", "correct")
  ) %>%
  select(-ACC, RT_log = logRT)

usethis::use_data(shallow, overwrite = TRUE)
```
