---
title: "07 - Linear models: Basics III"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Centring

---

- The estimated coefficients of the intercept and the effects of `vowel` are relative to when `c2_duration` is 0 ms.

- This is not very meaningful, because consonants cannot be 0 ms long (they would simply not be there!).

- We can get more meaningful coefficients by applying a type of transformation to `c2_duration` before we fit the model.

- This type of transformation is called **centring**.

---

**Centring simply consists of subtracting the mean of the variable from all the values.**

```{r dur-ita-cent}
dur_ita <- dur_ita %>%
  mutate(
    c2_duration_c = c2_duration - mean(c2_duration)
  )

dur_ita %>% select(vowel, v1_duration, c2_duration, c2_duration_c)
```

???

Since the mean of `c2_duration` in the data is `r mean(dur_ita$c2_duration)`, `c2_duration_c = 0` corresponds to `c2_duration = 90.33`.

The values of `c2_duration_c` are the difference between the value of `c2_duration` and the mean.

---

```{r c2-dur}
dur_ita %>%
  ggplot(aes(c2_duration)) + geom_density()
```

???

This is the density plot of `c2_duration`.

---

```{r c2-dur-c}
dur_ita %>%
  ggplot(aes(c2_duration_c)) + geom_density()
```

???

Here we plot `c2_duration_c`.

You can notice that the shape of the density is the same as before, but now the *x*-axis is different.

---

```{r vdur-lm-2-c}
vdur_lm_2_c <- lm(v2_duration ~ c2_duration_c * vowel, data = dur_ita)

tidy(vdur_lm_2_c)
```

???

Let's fit the model again, but now using `c2_duration_c`.

Now the estimates of the intercept and of vowel = /o/ and vowel = /u/ are interpreted as the estimates when `c2_duration_c` is 0, meaning when `c2_duration` is at its mean, i.e. `r mean(dur_ita$c2_duration)`.

---

```{r vdur-lm-2-c-pred-plot}
ggpredict(vdur_lm_2_c, terms = c("c2_duration_c", "vowel")) %>% plot()
```

???

The overall results are the same as in the model without centring.

What differs is just that the values of `c2_duration_c` now indicate the difference from the mean.

Centring is used mostly to make the interpretation of the coefficient in a model more meaningful.


---

layout: false
class: center middle inverse




