---
title: "SQM 02 - Data visualisation"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Statistics and Quantitative Methods (Frequentist) Tutorial 02: Dataviz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(learnr)
# tutorial_options(
#   exercise.blank = "..."
# )
library(glottologR)
data("glot_aes")
glot_aes$AES <- factor(glot_aes$AES, levels = c("not_endangered", "threatened", "shifting", "moribund", "nearly_extinct", "extinct"))
library(sqmf)
data("polite")
data("mald_1_1")
data("gestures")
```

## The basics

### Start here

The tidyverse package [ggplot2](https://ggplot2.tidyverse.org) provides users with a consistent set of functions to create captivating graphics. 
To be able to use the functions in a package, you first need to **attach the package**.
Here, we can attach the entire tidyverse, which includes ggplot2.

You only have to attach packages once, when you start your R session in RStudio.
They will stay attached until you close RStudio and end the R session.

Attach the tidyverse now.

```{r tidy-load, exercise=TRUE}
library(tidyverse)
```

### A basic plot

These are the minimum constituents of a ggplot.

- The `data`: you have to specify the data frame with the data you want to plot.

- The `mapping`: the mapping tells ggplot how to map data to parts of the plot like the axes or groupings within the data. These parts are called *aesthetics*, or `aes` for short.

You can specify the data and mapping with the `data` and `mapping` arguments of the `ggplot()` function.

Note that the `mapping` argument is always specified with `aes()`.

In the following bare plot, we are just mapping the mean f0 (`f0`) to the *x*-axis and the H2-H1 difference to the *y*-axis, from the `polite` data frame.

```{r bare-plot, exercise=TRUE, exercise.setup = "tidy-load"}
ggplot(
  data = polite,
  mapping = aes(x = f0mn, y = H1H2)
)
```

Not much to see here: just two axes!

### Let's add geometries

Nice, but we are missing the most important part: showing the data!

Data is represented with **geometries**, or `geom`s for short.
`geom`s are added to the base ggplot with functions whose names all start with `geom_`.

For this plot, you want to use `geom_point()`.
This geom simply adds point to the plot based on the data in the `polite` data frame.

Add the geom to the following code and run it.


```{r first-plot, exercise=TRUE, exercise.setup = "tidy-load"}
ggplot(
  data = polite,
  mapping = aes(x = f0mn, y = H1H2)
) +
  ...
```

This type of plot, with two continuous axes and data represented by points, is called a **scatter plot**

### Function arguments

Note that the `data` and `mapping` arguments don't have to be named explicitly (with `data =` and `mapping =`) in the `ggplot()` function, since they are obligatory and they are specified in that order

```{r named-arguments-1, exercise=TRUE, exercise.setup = "tidy-load"}
ggplot(
  polite,
  aes(x = f0mn, y = H1H2)
) +
  geom_point()
```

In fact, you can also leave out `x =` and `y =`.

```{r named-arguments-2, exercise=TRUE, exercise.setup = "tidy-load"}
ggplot(
  polite,
  aes(f0mn, H1H2)
) +
  geom_point()
```

Try running `?ggplot` in the Console to see the arguments of the function and the order they appear in.

```{r Q-named-arguments, echo=FALSE}
question(
  "Which of the following will produce the same plot as the one above?",
  answer('`ggplot(polite, aes(H1H2, f0mn)) + geom_point()`'),
  answer('`ggplot(polite, aes(y = H1H2, x = f0mn)) + geom_point()`', correct = TRUE),
  answer('`ggplot(polite, aes(y = f0mn, x = H1H2)) + geom_point()`'),
  correct = "Correct! Since the aesthetics are named, it does not matter that *y* comes before *x*.",
  incorrect = "Nope! The mapping of variable to aesthetics *x* and *y* is different from the plot above."
)
```







## Bar charts

Another common type of plot is the **bar chart**.

**Bar charts are useful when you are counting things.**
In the following example, we will be counting the number of languages by their endangerment status.

### Do you understand me?

There are thousands of languages in the world, but most of them are loosing speakers, and some are already no longer spoken.
The endangerment status of a language in the data is on a scale from `not endangered` (languages with large populations of speakers) through `threatened`, `shifting` and `nearly extinct`, to `extinct` (languages that have no living speakers left).

The `glot_aes` data frame contains the endangerment status for 7,845 languages from [Glottolog](https://glottolog.org).
Here's what it looks like.

```{r aes}
glot_aes
```

### Create a bar chart

To create a bar chart, add `geom_bar()` to the plot.

You only need one axis, the *x* axis to be precise, because the *y*-axis will have the counts.

```{r aes-bars, exercise=TRUE}
glot_aes %>%
  ggplot(aes(x = AES)) +
  ...
```

Note how the counting is done automatically.
R looks in the `AES` column and counts how many times each value in the column occurs in the data frame.

If you are baffled by that `%>%`, keep on reading.

### What the pipe?!

Wait, what is that thing, `%>%`?

It's called a **pipe**.
Think of a pipe as a teleporter.

In the previous code chunk, instead of specifying the data frame *inside* `ggplot()`, we teleport it into `ggplot()` by using the pipe.

So the following lines of code are equivalent.

```{r aes-pipe, eval=FALSE}
# Using the pipe
glot_aes %>%
  ggplot(aes(AES)) +
  geom_bar()

# Without the pipe
glot_aes(
  glot_aes, aes(AES)
) +
  geom_bar()
```

Don't worry too much if the concept is not clear yet.
It should become clearer in later tutorials.

### Stacked bar charts

A special type of bar charts are the so-called **stacked bar charts**.
These are bar charts where each bar contains a "stack" of shorter bars, each indicating the counts of some sub-groups.

To create a stacked bar chart, you just need to add a new aesthetic mapping: `fill`.
The `fill` aesthetic let's you fill bars or areas with different colours depending on the values of a specified column.

Let's make a plot on language endangerment by macroarea.

Complete the following code by specifying that `fill` should be based on `AES`.

```{r aes-stacked, exercise=TRUE, exercise.setup = "tidy-load"}
ggplot(
  data = glot_aes,
  mapping = aes(x = Macroarea, ...)
) +
  geom_bar()
```

```{r aes-stacked-hint}
aes(x = Macroarea, fill = ...)
```


What if we want to move the colour legend to the bottom of the plot?

Check out the documentation of `theme` by typing `?theme` in the RStudio Console and press ENTER.
Search for the word `position`.

```{r Q-legend-pos, echo=FALSE}
question(
  "Which of the following moves the legend to the bottom of the plot?",
  answer('`legend("bottom")`'),
  answer('`theme(legend.position = "bottom")`', correct = TRUE),
  answer('`theme(legend.bottom)`')
)
```

## Legends and labels

### Positioning the legend

The position of the legend can be set with the `legend.position` argument in `theme()`.

Move the legend to the `"bottom"`.

```{r aes-legend, exercise=TRUE, exercise.setup = "tidy-load"}
ggplot(
  data = glot_aes,
  mapping = aes(x = Macroarea, fill = AES)
) +
  geom_bar() +
  ...
```

```{r aes-legend-hint}
theme(legend.position = ...)
```

### Plot titles and labels

The `labs()` function allows you to set the plot title, subtitle, axis labels and more.

Since these are strings (i.e. text), they must be quoted with `"`.

```{r aes-title, exercise=TRUE, exercise.setup = "tidy-load"}
ggplot(
  data = glot_aes,
  mapping = aes(x = Macroarea, fill = AES)
) +
  geom_bar() +
  labs(
    title = "Number of languages by macro-area and\nAggregated Endangerement Status"
  )
```

NOTE: We use `\n` to split the title into two lines, so that it fits the plot. `\n` simply means "**n**ew line".

### Axis labels

To change the axis labels, you specify a string for `x` and `y`.

I have changed the `x` label here below.
Go ahead and change `y` to `"Number of languages"`.

```{r aes-labs, exercise=TRUE, exercise.setup = "tidy-load"}
ggplot(
  data = glot_aes,
  mapping = aes(x = Macroarea, fill = AES)
) +
  geom_bar() +
  labs(
    # We use \n to split the title into two lines. \n means "new line"
    title = "Number of languages by macro-area and\nAggregated Endangerement Status",
    x = "Language macro-area",
    ...
  )
```







## Density and violin plots

**Density plots show the distribution of the values of a continuous variable.**

Take for example the Reaction Times (`RT`) values in the *MAssive Lexical Decision* data set (`mald_1_1`).

```{r mald}
mald_1_1
```

### Reaction times

To plot the density of a continuous variable, you can use the `density` geometry.

Fill in the following code to create a density plot.

```{r mald-dens, exercise=TRUE, exercise.setup = "tidy-load"}
mald_1_1 %>%
  ggplot(aes(RT)) +
  ...
```

<div id="mald-dens-hint">
Remember that all geometries start with `geom_`?
</div>

### Make things cozy with a rug

The density line shows you a smoothed representation of the data distribution, but you might also want to see the raw data.

You can do so by adding the `rug` geometry.

```{r mald-dens-rug, exercise=TRUE, exercise.setup = "tidy-load"}
mald_1_1 %>%
  ggplot(aes(RT)) +
  geom_density() +
  ...
```


### Violin plots

An efficient way of showing the distribution of continuous variables depending on discrete groups (like `attitude`) are the so-called **violin plots**.

Go ahead and run the following code.

```{r pol-viol, exercise=TRUE}
polite %>%
  ggplot(aes(attitude, f0mn)) +
  geom_violin()
```

They are called violin plots because they look like violins!

### Lay over the raw data

You can lay over raw data on top of the violins.

To do so, you can use the `jitter` geometry.
Fill the following code chunk and run it.

```{r pol-viol-jit, exercise=TRUE}
polite %>%
  ggplot(aes(attitude, f0mn)) +
  geom_violin() +
  ...
```

That doesn't look great right?
The points are spread too widely.

We can fix that by specifying the argument `width` in the jitter geometry something smaller than `0.5`.
Try a few values until you think it looks better.

```{r pol-viol-jit-2, exercise=TRUE}
polite %>%
  ggplot(aes(attitude, f0mn)) +
  geom_violin() +
  ...
```

```{r pol-viol-jit-2-hint}
geom_jitter(width = ...)
```

Note that the order between the violin and jitter geoms is important.
If you add the jitter geometry before the violin geometry, the violins will just cover up the jittered points (see why it is called "layered" grammar of graphics?).


### Let's add some colour

Let's **fill** the violins with colour.
For now, just colour the violins depending on `attitude`

```{r pol-colour, exercise=TRUE}
polite %>%
  ggplot(aes(attitude, f0mn, ...)) +
  geom_violin()
```

```{r pol-colour-hint}
fill = ...
```

Not bad!
But maybe we can improve by making the fill a bit transparent?

You can achieve that with the `alpha` argument in `geom_violin()`, which takes a number between 0 (completely transparent) and 1 (completely opaque).

Add it to the code above to make the fill transparent.
Try different values until you are satisfied with the output.

### Separate the data by gender

In the previous section we have filled the violins with colours dependingn on `attitude`.
This helps making the plot somewhat more appealing, but colour there doesn't add information to the plot: attitude is already marked by the *x*-axis, so filling by attitude is a bit redundant.

Instead, we can use the fill argument to separate the data into further groups based on other variables in the data.

Go ahead and separate the data by `gender` using the `fill` argument.

```{r pol-colour-gen, exercise=TRUE}
polite %>%
  ggplot(aes(attitude, f0mn, ...)) +
  geom_violin()
```
