---
title: "SQM 03 - Transform data"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: Transform (filter, mutate, select) data with the tidyverse.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(learnr)
library(sqmf)
library(tidyverse)
shallow <- read_csv("https://raw.githubusercontent.com/stefanocoretta/quick-linear-models/main/04-fit-models/shallow-morph.csv")
```

## The data

For this tutorial, we will use the data from this paper: <https://doi.org/10.1017/S0272263120000170>.
Read the abstract for an overview. For more information, read the introduction and methods sections.

Here's the raw data (I have imported the data for you. You will learn how to import data in Week 4).

```{r shallow}
shallow
```

Remember that data tables (i.e. spreadsheet-like data) are called **data frames** or **tibbles** in R.

## Filter

**Data transformation** is a fundamental aspect of data analysis.

After the data you need to use is imported into R, you will very frequently need to filter rows, create new columns, or select some of the existing columns, among many other transformation operations.


### Filter data

You can filter a data frame with `filter()`, from the dplyr package (part of the tidyverse).

`filter()` needs two arguments:

- The data frame you want to filter.

- At least one rule to filter the data with.

The column `Critical_Filler` indicates which trials are the critical (experimental) trials or just fillers (distractors).
Let's filter `shallow` so that it returns only the critical trials (`Critical`).

Couldn't be easier!

```{r shal-filt, exercise=TRUE}
filter(
  shallow,                         # The data frame
  Critical_Filler == "Critical"    # The rule
)
```

The code above filters the rows of the `shallow` data frame so that only the rows where `Critical_Filler` equals (`==`) `"Critical"` are returned.

```{r Q-n-crit, echo=FALSE}
question(
  "How many critical trials are there in the tibble?",
  answer("6500"),
  answer("1950", correct = TRUE),
  answer("2000"),
  answer("11")
)
```

<div id="n-crit-hint">
When printing tibbles in the Console, the first line indicates the number of **rows** × **columns**.
</div>

Let's try again now using the pipe `%>%`.
Do you remember how it works?
Change the code below so that it uses the pipe.

```{r shal-pipe, exercise=TRUE}
filter(
  shallow,                         # The data frame
  Critical_Filler == "Critical"    # The rule
)
```

### Logical operators

The double equal sign we used above `==` is called a *logical operator*.
Logical operators are special symbols that allow you to compare objects.

These are the basic logical operators:

- `A == B`: A is *equal* to B.
- `A != B`: A in *not* equal to B.
- `A > B`: A is *greater* than B.
- `A < B`: A is *smaller* than B.

Logical operators return a "logical", i.e. either `TRUE` or `FALSE`.

```{r logic}
10 > 4
2 == 3
```

Try the logical operators out! Fill in the following code so that you get TRUE in the first 3 comparisons and FALSE in the last 3.
Substitute the `...` with numbers, strings, operators so that the statements return `TRUE` or `FALSE`

```{r logic-ex, exercise=TRUE, eval=FALSE}
# Use numbers or strings.
# These should return `TRUE`
... < 20
"a" == ...
"apple" != ...

# Use operators.
# These should return `FALSE`
"zoo" ... "park"
120 ... 43
0.1 ... 0.002
```

### Multiple filtering

You can filter a data frame based on more than one column.
Simply add more rules, separated by commas.

```{r multi-filt, exercise=TRUE}
shallow %>%
  filter(
    Critical_Filler == "Critical",
    RT > 1000
  )
```

```{r Q-multi-filt, echo=FALSE}
question(
  "Which of the following statements is correct?",
  answer("There are 10 observations that are critical trials and in which the reaction times are greater than 1000 ms."),
  answer("Out of a total of 6500 observations, 616 are critical trials or have reaction times that are greater than 1 s."),
  answer("There are 616/6500 trials that are critical and generated reaction times that are greater than 1 s.", correct = TRUE)
)
```

Now filter the data so that you get only those trials that are critical, unrelated relation type, right branching and have RTs that are lower or equal to 494 ms.

```{r shal-filt-2, exercise=TRUE}
shallow %>%
  filter(
    ...
  )
```

<div id="shal-filt-2-hint">
The less or equal and more or equal operators are `<=` and `>=` respectively.
</div>


## Mutate

What if you want to create a new column based on existing columns?

You can create new columns with the `mutate()` function from dplyr.

### How to create a new column

`mutate()` needs the name of the data frame you want to mutate, the name of the new column and the code to create it.

```{r long-rt, exercise=TRUE}
shallow <- shallow %>%     # Ceci c'est un pipe. ;)
  mutate(
    long_RT = RT >= 1000
  )

shallow
```

Woah! A lot of new stuff going on.
Let's unpack it.

This time we have saved the output of `mutate()` back into `shallow` using the assignment symbol `<-`, so that we can use the new column in further code.
In other words, we have overwritten `shallow` with the mutated data frame.

Then we use `mutate()` to create a new column called `long_RT`.

Note that the name of the new column is followed by *one* equal sign `=`, not two `==`.
We are not comparing objects, but creating one.

We are telling `mutate()` to fill the new column with `TRUE` if `RT` **is greater than or equal to** `1000` ms, with `FALSE` if `RT` **is not greater than or equal to** `1000` ms.

To test if `RT` is greater than or equal to `1000` ms, we are using the logical operator `>=`.
We are comparing two objects here.

### If, else?

Sometimes you will need to create columns based on multiple criteria.

There are several ways to do that, but a very common one is by using "if-else" statements.

If-else statements are exactly what you think they are: **if** X is true, do A; **else** if false, do B.

The following code will return `"short"` if the RTs are smaller than 1 s, otherwise it will return `"long"`.

```{r if-else, exercise=TRUE}
RT <- c(554, 1098, 1000, 245, 2502)

ifelse(RT < 1000, "short", "long")
```

Now we can use the `ifelse()` function to create a new column in `shallow`.

The new column, called `accuracy` should have `"incorrect"` if the `ACC` value is `0`, else it should be `"correct"`.
Fill in the following chunk with the right code.

```{r if-else-acc, exercise=TRUE}
shallow <- shallow %>%
  mutate(
    accuracy = ...
  )

shallow
```

<div id="if-else-acc-hint">
Remember, to check if two objects are **equal** you need the `==` operator.
</div>

Note that `>` and `<` mean "greater than" and "less than".
This means that `1000 < 1000` is `FALSE`.

The operators `>=` and `<=` mean "greater than or equal to" and "less than or equal to".
So `1000 <= 1000` returns `TRUE`.

### Count occurences

We can use the `count()` function to count the number of occurrences for each value of a column.
Let's count how many trials are correct (`accuracy`).

`count()` takes the name of the column you want to count values in.

```{r count-corr}
shallow <- shallow %>%
  mutate(
    accuracy = ifelse(ACC == 0, "incorrect", "correct")
  )

shallow %>%
  count(accuracy)
```

You can add multiple columns to get counts for the combinations of values of each column.

Try to get counts of accuracy and `Group`.

<div id="count-corr-hint">
In `count()`, include the names of the two columns you want to get counts of, separated by commas.
</div>


## Select

What if you only want some columns? Use the `select()` function.

`select()` takes a list of unquoted column names (i.e. the bare names, without quotes `"`) separated by commas.

```{r select, exercise=TRUE}
shallow %>%
  select(Group, Relation_type)
```

To remove one or more columns, prefix the column name with `-` (minus). All the other columns will be kept.

```{r deselect, exercise=TRUE, exercise.setup="mud-count-setup"}
shallow %>%
  select(-List, -RT)
```

To select a range of columns you can include only the first and the last of the range, separated by `:`.

```{r select-range, exercise=TRUE}
shallow %>%
  select(logRT:Branching)
```


### You did it!

Success! You completed this tutorial.

Now go through the tutorial `03b_lm_basics`!
