---
title: "SQM 03 - Linear Model Basics"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library(sqmf)
library(tidyverse)
data("mald_1_1")
data("polite")
data("dur_ita_pol")
```

## Does phoneme-level distance affect reaction times?

### The MALD data

We will use `mald_1_1` (made available through the sqmf package), which contains data from a lexical decision task.

We want to know if phoneme-level distance (`PhonLev`, i.e. Levenshtein distance) affect the reaction times of respondents when they need to decide whether the word they heard is a real word or a nonce word.

Here's what the data looks like.

```{r mald, exercise=TRUE}
mald_1_1
```

### Phoneme-level distance vs reaction times

Let's start by plotting phoneme-level distance (`PhonLev`) vs reaction times (`RT`).

You want to use a *scatter plot*.

```{r mald-plot, exercise=TRUE}
mald_1_1 %>%
  ...
```

<div id="mald-plot-hint">
A **scatter plot** is a plot with continuous variables in the *x* and *y* axes.
</div>

### Model distance vs RTs

Now we can run a simple linear model to investigate the relationship between phoneme-level distance and reaction times.

Fill in the following code with the right formula.

```{r mald-lm, exercise=TRUE}
lm_1 <- lm(..., data = mald_1_1)
```

And now look at the model summary!

```{r mald-summary-setup}
lm_1 <- lm(RT ~ PhonLev, data = mald_1_1)
```

```{r mald-summary, exercise=TRUE}
...
```

```{r mald-summary-hint}
summary(...)
```

### Does phoneme-level distance affect reaction times?

Great! Now that you have seen the model summary, answer the following questions.

```{r Q-lm-1-1, echo=FALSE}
question(
  "Which of the following statements is correct?",
  answer('The mean RT is 771.2 ms.'),
  answer('The model intercept is 33.7 ms.'),
  answer('When distance is 33.7, the average reaction time is 771.2 ms.'),
  answer('On average, the reaction time is 771.2 ms, when distance is 0.', correct = TRUE),
  correct = "Correct!",
  incorrect = "Nope!"
)
```

```{r Q-lm-1-2, echo=FALSE}
question(
  "The effect of `PhonLev` on `RT` is",
  answer('positive.', correct = TRUE),
  answer('negative.'),
  answer('zero.'),
  answer('3.4 ms.'),
  correct = "Correct! The sign of the effect is `+`, so the effect is positive.",
  incorrect = "Nope!"
)
```

```{r Q-lm-1-3, echo=FALSE}
question(
  "Which of the following statements is correct?",
  answer('When the intercept is 0, phoneme-level distance is 33.7.'),
  answer('For every unit increase of phoneme-level distance, there is an increase of 33.7 ms.', correct = TRUE),
  answer('When phoneme-level distance is 0, the reaction time is 33.7 ms.'),
  answer('For every 1 ms, the phoneme-level distance increases by 33.7.'),
  correct = "Correct!",
  incorrect = "Nope!"
)
```

```{r Q-lm-1-4, echo=FALSE}
question(
  "What is the effect of phoneme-level distance when the distance is 3?",
  answer('101.1 ms.', correct = TRUE),
  answer('0 ms.'),
  answer('33.7 ms.'),
  answer('3.4 ms.'),
  correct = "Correct! To get the effect when distance is 3, you just do `33.7 * 3`.",
  incorrect = "Nope!"
)
```
